The Astropy Project

### 2017-11-13 main.tex
ack line
list of new features for Table

\subsubsection{Tables}
\label{sec:table}

The \package{Tables} package provides wrappers to \package{NumPy} structure arrays, which can contain heterogeneous data sets.   The package was described in detail in \cite{astropy}. Tables now has the ability to work with other astropy classes including Quantities, Time, and Coordinates.  In addition, operations can also be performed on Tables.

%QTable is new, mixin columns for time and coordinates. Table operations were added in v0.3

@@

Software/services to thank: GitHub. Travis CI, Appveyor, CircleCI. Read the Docs.

%% Similar to \facility{}, there is the optional \software command to allow
%% authors a place to specify which programs were used during the creation of
%% the manusscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

\software{\package{astropy} (\citealt{astropy}),
          \package{numpy} (\citealt{numpy}),
          \package{scipy} (\citealt{scipy}),
          \package{matplotlib} (\citealt{matplotlib})
          \package{Cython}(\citealt{cython}),
          }

>>>>

\subsubsection{Tables}
\label{sec:table}
% fleshed out / edited by Adrian PW

The \package{astropy.table} subpackage provides functionality for representing
and manipulating heterogeneous data.
\inlinecomment{APW:}{Need more introduction to the table subpackage!}
The package was described in detail in \cite{astropy}; we summarize key new
features or updates to \package{astropy.table} below.

\begin{description}
    \item[Support for grouped table operations] Some tables contain data that
        naturally form groups.
        For example, a table of photometry for galaxies, stars, and quasars.
        We may then want to split the table or a column into these groups and
        compute some statistic on the data, or filter the groups based on some
        criteria.
        Grouped operations are now supported by \texttt{Table} objects,
        mainly through the \texttt{group\_by()} method and subsequent operations
        on the grouped version of the table or column.
    \item[Support for table concatenation] Tables can be combined in several
        different ways.
        If two tables have the same rows, we may want to stack them
        ``vertically'' to create a new table with the same columns but all rows.
        If two tables are row-matched but have distinct columns, we may want to
        stack them ``horizontally'' to create a new table with the same rows
        but all columns.
        For other situations, more general table concatenations or joins are
        also possible when two tables share some common columns.
        All three of these operations are now possible with \texttt{Table}
        objects through the \texttt{vstack}, \texttt{hstack}, and \texttt{join}
        functions in \package{astropy.table}.
    \item[Mixin columns for time and coordinates] The default data containers
        for \texttt{Table} objects are \texttt{Column} objects.
        However, sometimes the data for a given column is better represented as
        a \texttt{Quantity}, or a sky coordinate (\texttt{SkyCoord}), or an
        astronomical \texttt{Time} object.
        The \texttt{Table} object now supports these and other objects to serve
        as array-like containers for column data through the ``mixin'' protocol.
    \item[Transform to and from \texttt{Pandas} \texttt{DataFrame} objects]
        \texttt{Table} objects can now be created from a \texttt{DataFrame}
        object (\texttt{from\_pandas()}), or exported to a \texttt{DataFrame}
        object (\texttt{to\_pandas()}).
        \texttt{DataFrame} objects do not support multi-dimensional array
        columns, so this conversion will fail if the source \texttt{Table}
        object contains any such columns.
\end{description}

@@

We thank the many web services that make developing and maintaining this package
feasible, such as GitHub, Travis CI, Appveyor, CircleCI, and Read the Docs.

\software{\package{astropy} (\citealt{astropy}),
          \package{numpy} (\citealt{numpy}),
          \package{scipy} (\citealt{scipy}),
          \package{matplotlib} (\citealt{matplotlib}),
          \package{Cython}(\citealt{cython}),
          }
